### 2021-06-24 ch1intro/ch1intro.tex
gearing up for intense work.

%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Sediment transport and landscape evolution}
\label{ch:Introduction}

Landscapes evolve across geologial time due to energy and mass contributed from above and below the Earth's surface.

\section{Stochastic properties of sediment movement}

\section{Historical survey: statistical physics in sediment transport}

\section{Philosophy of the statistical physics approach}

\section{What's left?}

\endinput

Any text after an \endinput is ignored.
You could put scraps here or things in progress.