### 2021-06-24 ack.tex

%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Acknowledgments}

Thanks Leo Golubovic, Mindy Saunders, Calisa and Jim Pierce, Kim and Kelsey Pierce, Johnathan Ramey, Linda Mendez, Tex Wood, Charles Wood, and all of the Appalachian intelligensia who paved my way. Tyler cannon. . .

Shawn Chartrand, Conor McDowell, Matteo Saletti, Will Booker, David Adams, Leo King, Tobias Mueller, Nisreen Ghazi,

Adam Payne, Adam Collins, Megat Denney, Andrew Rice, Evan Graber, Scott Ferris, Craig Tenney, and the rest of the WVU physics lounge family.

Andy Osadetz, Josh Plankeel,

My partner Mary for putting up with the late nights.