### 2021-06-24 formerPapers/ens-rev/reviewsections/heyman2014.tex

\subsection{Spatial correlations in bedload transport: }

\citet{Heyman2014} used the \citet{Ancey2014} model to examine a set of experiments resolving spatial correlations of bedload transport. 
They found a stochastic differential equation for the mean of the poisson rate $\eta$: 
They related this to a point process framework which allowed them to analyze spatial correlations in the bedload rate and relate this to their birth-death model. 
Somehow this whole thing is contingent on their two-point spatial correlation process. 
They state that the whole problem is linear so that it'd be straightforward to extend it to 2d. 
slowing down ...