### 2021-06-25 ack.tex

%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Acknowledgments}
Thanks Leo Golubovic, Mindy Saunders, Calisa and Jim Pierce, Kim and Kelsey Pierce, Johnathan Ramey, Linda Mendez, Tex Wood, Charles Wood, and all of the Appalachian intelligensia who paved my way. Tyler cannon. . .

The people whose benevolence spans generations: both those I know: Joe, Joyce, Ama, Ivory Nichols, and Parry Steele, and those I was unlucky to miss. Maragret  Squire Avis and Oddessa Mathis. Joe, Joyce, Ama, and Parry Steele. My

Shawn Chartrand, Conor McDowell, Matteo Saletti, Will Booker, David Adams, Leo King, Tobias Mueller, Nisreen Ghazi,

Adam Payne, Adam Collins, Megat Denney, Andrew Rice, Evan Graber, Scott Ferris, Craig Tenney, and the rest of the WVU physics lounge family.

Andy Osadetz, Josh Plankeel,

My partner Mary for putting up with the late nights.

Mike church and olav slaymaker for the inspiring seminars.

marwan hassan for being one of the most emotionally intelligent people I've ever met.