### 2021-06-25 appendix.tex

%%!TEX root = diss.tex

\chapter{Mathematical Compendia}