### 2021-07-02 ack.tex
got another intro figure constructed.

%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Acknowledgments}

Attending university long enough to earn a PhD is an unbelieveable priviledge, and I only got here with an incredible amount of guidance and support.

Most of all, to mom, dad, Kelsey, Kim, and Gugs, you've formed all of the best parts of my personality, and it's impossible to explain how thankful I am for that.

Mom, during my childhood, you worked full time, taught piano lessons, earned Master and Doctoral degrees, played organ every Sunday, consoled us through every difficulty, and engaged us in as many extracurricular activities as we were willing to take on. 
I cannot imagine a better example of work ethic, open-mindedness, and devotion. I hope I picked a few things up.

Dad, I've watched you solve an insane range of problems without hesitation, from worn out brake pads to broken arms.
By age 10 I must have helped with copper plumbing, woodwork, car repair, home wiring, and tree felling.
I had personal tours of West Virginia's backroads and civil engineering works, and I learned more about science paddling the South Branch with you than I ever did in school.
To you, water currents, fish habits, electrons, hellgrammites, engines, heat, and radio waves are all deeply important, and any man-made problems are easy enough to solve with time and the right tools.
Thanks for sharing the perspective of a problem solver and knowledge seeker! And of course so much more.

Kelsey and Kim, my earliest memories are sitting in your bedroom floor struggling through your books, asking you for help every couple of paragraphs. You made me good at reading and taught me to enjoy it, which is the foundation of all my learning. Plus, we had a lot of hilarious times. I know it goes without saying, but thanks.

Gugs, your approach to life is clean and selfless, and I cannot imagine a single way you could have been more reliable, stable, or invested in my development.
I'll never be able to thank you enough!
One day I'll take up your 4am wakeups and adopt your completely measured and calm responses to life's challenges. Until then, I'll keep thinking about I'd ideally do in terms of what you always do, and if we're lucky I might even occasionally pull it off.

Another family member who deserves special mention is my great aunt Alice. Thanks so much for paying for my undergraduate education. We hardly met, but we're family, and that was enough. It really means a lot.

Thanks to my early educators who put in the extra care to keep me on track, Jonathan Ramey, Linda Afzhalirhad, Brandon Willard, Linda Mendez, Marty Ojeda, Phyllis Adkins, and Red Hamlin among many others.
Thanks also to the later ones at Southern WV Community College (SWVCC) and West Virginia University (WVU) whose lessons fostered my way of thinking and provided me with many of the tools I used for the research in this thesis. At SWVCC, I'd like to thank especially Mindy Saunders, Tex Wood, Anne Klein, Charles Wood, Don Saunders, and George Trimble. Mindy in particular transformed my view of the world into one based on mechanics, and provided me with a deep understanding of general physics and calculus. Everyone there always encouraged me to continue my education. You all made SWVCC a personal transformation and an unbelievable privilege.

At WVU, I'd like to thank Alan Bristow, Tudor Stanescu, and Leo Golubovic. Alan and Tudor, thanks for providing my first opportunities in research. Leo, you are the key person in my intellectual development. You inspired me to learn mechanics, statistical physics, and mathematical physics at a high level, and I can't thank you enough. Every time I solve a challenging problem with some Bessel functions or contour integrals, I think ``Dr. Golubovic would like this one." I wish I had taken more advantage of the opportunity to speak to you outside of classes. I'll feel deeply successful if I can one day become as inspiring an educator as you are.

To my extended cohort at WVU physics, Payne, Megat, Collins, Scott, Dustin, Andrew Rice, Andrew Wolfe, Samet, Craig, April, Evan, Larry, Robert, Stephen, Rick, Matt, Caleb, John, and all of you others, thanks for the collaboration and the hilarious times. I'll never forget Scott's quaternion obsession, Samet's olympiad problems, or Megat's power tower magic (which still blows my mind). I have yet to experience teamwork and shared exploration like we had in the undergrad lounge. I hope the lounge is still happening, that its residents by some miracle still have 24 hour access into the building, that some barefoot Megat-like character is still gaming in the corner, and that they have some Andrew Rice imitator to flip some cars for em (mike jones). Because of you all, WVU was magic for me.

Now we come to Marwan Hassan. About five years ago now I wandered into Marwan's office saying I liked rivers and wanted to describe sediment movement with statistical physics. I realize now when I said ``statistical physics" your eyes lit up with thoughts of Einstein, Nakagawa, Yano, Furbish, Hubbell, and all of the names I now know well. I was lucky to use the right trigger words.
Marwan, thanks for taking the risk and taking me on from outside the field, for all of the guidance, for loaning me so many papers and books, giving me freedom to explore ideas, and most of all for showing me that it's ok to be human in science. At this point I also want to thank Yusuf for trying your patience to its limits over a long enough period so that things didn't seem so bad once I came along for this short time together.

My friends in geography have also been great. Thanks to the whole extended research family: Shawn, Matteo, Conor, Nisreen, Yinlue, Alex, Kyle, Leo, Dave, Tobias, Katie, Maria, Elli, David, Jiamei, Xingyu, Niannian, and Emma, plus the outsiders, Will, Dave, Rose, and Anya. In particular, thanks to Shawn for all of the chats and shared speculations, and Matteo and Conor -- friendships forged through suffering! We'll have to keep in touch.

Thank you Rui Ferreira and Brett Eaton!
Rui, you welcomed me to Portugal, offered me guidance in the laboratory, and educated me on turbulence and hydraulics. Your kindness and wide-ranging interests beyond hydraulics are inspiring. Thanks for the thoughtful comments on my PhD. I'm not forgetting what I owe you! 
Brett, you were the first person to welcome me to UBC, and I'll always appreciate your effort to teach me how to write, interpret flume experiments, and understand modelling in light of the real world and its complexity. Although I'm not sure if I understood your lessons as well as you would have liked, I have a list of your favorite books on my computer, and they're a first task for as soon as I defend. Maybe they'll bring me closer to your way of thinking.

I'd be amiss not to mention David Furbish. Before I started my PhD, when I was still exclusively a physics kid, I attended Furbish's colloquium in Geography at Marwan's invitation. At the end I asked a vague question about whether Langevin equations might be useful for describing sediment transport, and the answer was ``absolutely yes". Now I have a thesis on it. Later on, David recruited me into a lovely series of meetings where we discussed gases with adhesion, moon craters, entropy, soft matter, martian hillslopes, the social repercussions of disinformation, and a whole range of questions including the meaning of probability in landscape evolution, probability as being just as fundamental as mass, momentum, and energy, and fringe ideas in cosmology. These are my people! David, Rachel, Sarah W, Sarah Z, Nakul, Tyler, Erika, and Shawn, thanks sincerely for letting me hop on the spacecraft.

Among UBC Geographers, thanks especially to Nina Hewitt! It was always a pleasure and maybe a sigh of relief to TA your courses, which are so uniquely organized and student-focused. UBC Geography is a much greater department with you in it. I'm so thankful to have been associated with your influence on the undergrads. I'll be happy if I can teach like you one day!

UBC Geography is successful because it has excellent staff, Vincent Kujala, Sandy Lapsky, Bret Peterson, Suzanne Lawrence, Connie Cheung, Jose Aparicio, thank you all so much! 
Vincent showed me the magic of bash, Sandy handled conference expenses and got me back in an office during the pandemic when my productivity was floundering, Bret put out the tech fires, Suzanne organized the paychecks and appointments, and Jose handled every online teaching challenge. You all saved me a ton of time, and I really appreciate it.

The kind, open-minded, and collaborative atmosphere I experienced here is clearly a legacy from the former generation of UBC's physical geographers, some whom I had the pleasure of meeting and learning from, like Mike Church and Olav Slaymaker, and others whom I didn't. Thanks Mike and Olav for your inspiring seminars and the academic environment you cultivated.

Finally, Mary! My PhD has not been great in terms of work-life balance. Thanks for sticking it out, putting up with my late bedtimes, my early wakeups, and all of those social events I didn't come to. It's time! I'm done. The pandemic also seems to be winding down. We can finally get to Aus! I'm ready for an extensive tour of the gap and some beautiful Queensland beaches with your family, featuring SPF110 and an incredibly giant umbrella. After that, we're a step closer to settling somewhere and determining once and for all whether our future cat will be white or orange.

Completing a PhD looks like a personal achievement, but it's really not.
It's an expression of 30 years of influence and guidance from other people at every stage.
I'm kinda tired because it's the end. Maybe because of it, I find it easy to drift into daydreams on the parts of my life I've really loved, Watoga and Williams River with my family, reading encyclopedias in Gugs' living room, lining up just left of center with my dad and bracing on the paddle for the hit, nighttime fireflies under my great grandpa's tree, and sliding down rocks at Kumbrabow with my sisters. The daydream shifts to years later, as I flip on the light in the physics lounge and surprise Megat who was asleep on the couch, I finish a problem set with friends, discuss Tsujimoto with Marwan (``those guys did everything!"), or just walk around Kitsilano with Mary at the edge of dark. It's all one thing! Thank you all for every part of it.