### 2021-07-02 app_ch2.tex
got symbols working in the toc.

%%!TEX root = diss.tex

\chapter{Ch. 2 calculations}

\section{Derivation of the master equation}
\section{Solution for the position probability distribution}
\section{Moments of position}
\section{Calculation of the flux rate constant}