### 2021-07-03 ch6conc.tex
got all references fixed up... except for a few which were overwritten by mendeley like taylor 1920. Still need to check carefully what's actually cited.

%%!TEX root = ../diss.tex

\chapter{Summary and future work}
\label{ch:conc}
\section{The overall strategy}

This thesis has presented a review and four chapters which generalize and extend earlier work to describe sediment transport as a stochastic process. In the research, I characterized the movements of grains along a wider range of timescales than before, and I developed new a new formalism using these movements to predict the sediment flux.

\subsection{Langevin and master equations}

Probability

\subsection{Idealized noises and their combinations}

\section{Key contributions}

\subsection{Probability distribution of the sediment flux from micromechanics of particle trajectories}

\subsection{Inclusion of velocity fluctuations into Einstein's model of individual particle trajectories}

\subsection{Quantification of the control of bed elevation fluctuations over sediment transport fluctuations}

\subsection{Predicting how sediment burial affects the downstream spreading of sediment tracer particles}

\section{Models and the real world}

\section{Conclusion and next steps}