### 2021-07-04 laySum.tex

%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

%% https://www.grad.ubc.ca/current-students/dissertation-thesis-preparation/preliminary-pages
%% 
%% LAY SUMMARY Effective May 2017, all theses and dissertations must
%% include a lay summary.  The lay or public summary explains the key
%% goals and contributions of the research/scholarly work in terms that
%% can be understood by the general public. It must not exceed 150
%% words in length.

\chapter{Lay Summary}

Predicting the movement of gravel through river channels is important to manage environmental issues involving rivers, since moving gravel can destabilize river channels.
A majority of models assume gravel flows as if it were a heavy fluid, but this description is crude considering how gravel actually moves.
In this thesis, I instead describe gravel movement as the result of individual grains bouncing and rolling along the streambed.
Using this concept, I calculate the downstream flow rate of gravel, study how this flow rate depends on the shape of the bed, and predict how quickly individual grains move downstream.
This research produces a more realistic description of how gravel moves in river channels which helps us understand how rivers change through time.

\endinput
The lay or public summary explains the key goals and contributions of
the research\slash{}scholarly work in terms that can be understood by the
general public. It must not exceed 150 words in length.