### 2021-07-04 app_ch2.tex
rolling on the conclusion

%%!TEX root = diss.tex

\chapter{Calculations involved in dynamical sediment flux model}

\section{Derivation of the master equation}
\section{Solution for the position probability distribution}
\section{Moments of position}
\section{Calculation of the flux rate constant}