### 2021-09-03 revision/formerPapers/ens-rev/scraps/mechanics.tex

\section{Underlying mechanics of Markov models: toward the mechanistic-stochastic} 
\subsection*{Mechanics versus kinematics} 
words

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Mechanical perspectives on the entrainment probability} 
words

Entrainment probability initially considered by \citet{Einstein1942, Einstein1950}. 
\citet{Grass1970} is very important. 
Many authors followed in the Grass paradigm to define particle entrainment as a convolution integral over a turbulent fluctuation pdf \citep{Paintal1971, Kirchener1990, Ling1995, Laursen1999, Wu2002b, Wu2004a, Wu2003, Papanicolaou2002, Bridge1992, Kirchener1990, Ma2013, Nakagawa1980, Vanrijn1984}
These approaches have been somewhat undermined by the recent developments in the impulse idea \citep{Diplas2008, Valyrakis2011, Valyrakis2013, Celik2014}. 
Investigators have just begun to bring these two research directions into accord with one another \citep{Tregnaghi2012, Dey2018}.

\subsection*{Mechanical perspectives on particle travel characteristics}

hydrodynamics of the particle velocity distribution: 
random walks to describe the distribution: \citep{Ancey2008, Fan2014}
velocity ensembles, \citep{Furbish2012b}, \citep{Furbish2013} 
what controls the particle hop distance distribution? 
Furbish something? where? \citep{Furbish2016}

\citet{Martin2014}: Distribution of waiting times from a mean reverting random walk model

\subsection*{Mechanics of sediment deposition} 
\citep{Charru2004} is a kinematic model. 
There is the recent paper on thresholds of deposition
See what else you can find.

\subsection*{Analyzing assumptions: Are kinematic models enough?}
Houssais argues that bedload transport is a continuous transition from granular creep which would undermine the
partitioning scheme into motion and stationary states.
Possible that multiple motion modes exist meaning more states necessary in markov models.