### 2021-09-03 revision/laySum.tex

%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

%% https://www.grad.ubc.ca/current-students/dissertation-thesis-preparation/preliminary-pages
%% 
%% LAY SUMMARY Effective May 2017, all theses and dissertations must
%% include a lay summary.  The lay or public summary explains the key
%% goals and contributions of the research/scholarly work in terms that
%% can be understood by the general public. It must not exceed 150
%% words in length.

\chapter{Lay Summary}

Predicting the flow rates of gravel through river channels is important to manage environmental issues involving rivers.
Most approaches to predict gravel flow consider gravel as if it were a heavy fluid.
This description oversimplifies the problem, since gravel flow is actually individual grains bouncing and rolling downstream, which is hardly like a fluid.

In this thesis, I consider gravel flow as the result of individual grains. Using physics and probability, I calculate how fast individual grains move downstream, then I demonstrate how to use their downstream movement paths to predict overall gravel flow rates.
This research produces a more realistic description of gravel movement in rivers which helps us understand how rivers change through time.

\endinput
The lay or public summary explains the key goals and contributions of
the research\slash{}scholarly work in terms that can be understood by the
general public. It must not exceed 150 words in length.