### 2021-09-23 revision/ack.tex
uh

%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Acknowledgments}

Attending university long enough to earn a PhD is an unbelievable privilege, and I only had access to it thanks to an incredible amount of guidance and support.

Most of all, to mom Calisa Pierce, dad Jim Pierce, my sisters Kelsey and Kim, and my grandmother Wilma Avis, it's impossible to explain how thankful I am for your guidance. Thanks also to my great aunt Alice who graciously paid for my undergraduate education.

Thanks to my early teachers who put in the extra care to keep me on track, Jonathan Ramey, Linda Ball, Damon Spurlock, Linda Mendez, Linda Afzalirad, Phyllis Adkins, Brandon Willard, Marty Ojeda, and Red Hamlin among many others.

Thanks also to my professors at Southern WV Community College who fostered my way of thinking and provided many of the tools I used for the research in this thesis.
At SWVCC, I'd like to thank especially Mindy Saunders, Tex Wood, Anne Klein, Charles Wood, Don Saunders, and George Trimble. SWVCC was a personal transformation and a great privilege.

At WVU, I'd like to thank Alan Bristow, Tudor Stanescu, and Leo Golubovic in particular. Alan and Tudor, thank you for providing my first opportunities in research. Leo, you are the key person in my development as a researcher.

To my extended cohort at WVU physics, Payne, Megat, Collins, Scott, Dustin, Rice, Wolfe, Samet, Craig, April, Evan, Larry, Robert, Stephen, Rick, Matt, Caleb, John, and all of you others, thanks for the collaboration and the hilarious times. May the lounge live on forever!

To Marwan Hassan, some time ago now I wandered into your office saying I liked rivers and wanted to describe sediment transport with statistical physics. I realize now when I said ``statistical physics" your eyes lit up with visions of Einstein, Nakagawa, Yano, and so many other names I now know well. I was lucky to use the right trigger words.
Marwan, thanks for all of the guidance, for lending me hundreds of books and papers, for the freedom to explore ideas, and most of all for showing me that it's ok (and implicitly expected) to be human in science.

My friends in geography have also been great. Thanks to the whole extended research family: Shawn, Matteo, Conor, Nisreen, Yinlue, Alex, Kyle, Leo, Dave, Tobias, Katie, Maria, Elli, Jiamei, Xingyu, Niannian, and Emma, plus the outsider Eaton-ites, Will, Dave, Rose, and Anya. In particular thanks to Shawn for all of the shared speculations.

Thank you Rui Ferreira and Brett Eaton!
Rui, you welcomed me to Portugal, offered me guidance in the laboratory, and educated me on turbulence and hydraulics. Your kindness, physics knowledge, and wide-ranging interests beyond physical science are inspiring. Thanks for the thoughtful comments and advice on my PhD.
Brett, you were the first person to welcome me to UBC, and I'll always appreciate your effort to teach me how to write, interpret flume experiments, and understand simplified modeling in light of real world complexity.

To David Furbish, before I started my PhD when I was still a physics kid exclusively, I recall attending your colloquium at UBC with Marwan's invitation. At the end I asked a vague question about whether Langevin equations might be useful for describing sediment transport, and the answer was ``absolutely yes". Now I have a thesis on it. Thanks for the confirmation! And to the Furbish people -- Rachel, Sarah W, Sarah Z, Nakul, Tyler, Erika, and Shawn, thanks sincerely for letting me board the spaceship.

Among UBC Geographers I'd like to thank Nina Hewitt in particular. It was always a pleasure to work for you on your courses, which are so uniquely organized and student-focused. UBC Geography is a much greater department with you in it.

Finally, Mary! My PhD has not been great in terms of work-life balance. Thanks for putting up with my late bedtimes, early wakeups, rambling science jargon, and skipped social events. It's done!

Completing a PhD looks like a personal achievement, but it's really not.
It's an expression of 30 years of influence and guidance from other people at every stage, my parents, my sisters, my grandmother, and a huge network of teachers and mentors. Thank you all!

\endinput

Mom, during my childhood, you worked full time, taught piano lessons, earned Master and Doctoral degrees, played organ every Sunday, consoled us through every difficulty, and engaged us in as many extracurricular activities as we were willing to take on.
I cannot imagine a better role model of work ethic, open-mindedness, and devotion. I hope I picked a few things up!

Dad, in my childhood I watched you solve an insane diversity of problems without hesitation, from easy ones, like worn out brake pads or stuck chainsaws, to hard ones, like burnt eyes and broken arms.
On canoeing trips I got lessons in water currents, fish habits, tree species, electrons, engines, wings, hellgrammites, heat transfer, world history, radio waves, and probably everything else I could think of. Years of free science lessons! You taught me more about science and problem solving than school ever did. "Any problem can be fixed with the right tools" is basically the Jim Pierce philosophy. It it's incredibly useful in a PhD! Thanks for this and so much more.

Kelsey and Kim, my earliest memories are sitting in your bedroom floor struggling through your books, asking for help every couple of paragraphs. You made me good at reading and taught me to enjoy it, and this is the basis of everything. Kim, you were my first role model, and I think of you all the time. Kelsey, I know it goes without saying, but thanks for a lifetime of guidance. You're now, as always (all state chorus, math field day, spelling bees, piano skills, grades, . . . ), five steps ahead and far more successful. On the one hand, I'm deeply proud at your progress through life, with a stable career and now two beautiful kids, but on the other, I'm flatly shocked by it.
I'll be playing catch up forever!

Gugs, you're easily the most level-headed, good-natured, and reliable person I've ever had contact with, and to think I was lucky enough to be about half raised by you! You've been my second mother, my counsellor, at times my daily chauffeur, and my best role model of grace, consistency, and even-handedness. I'll never be able to thank you enough! One day I'll finally take on your 5am wakeups and figure out how to copycat your completely measured responses to all of life's challenges.

Finally with regard to family, I'd like to thank my great aunt Alice. Thanks so much for paying my undergrad education! We hardly met before you passed, but we're family, and that was enough. It really means a lot.

Next comes the mentors in education.

I'm kinda tired because it's the end, so it's natural to drift on to recollections of my favorite times in life, Watoga and Williams River with my family, encyclopedias in Gugs' living room, watching eagles over the South Branch with dad, discussions with Johnathan Ramey about my education, the glowing fireflies under my great grandpa's tree, ``cross-up, cross-up, squeeze" in Mindy's calc class, or laughter about stolen watermelons with my sisters. I can skip years later and flip on the light in the physics lounge to find Megat asleep on the couch, visit Gugs on a trip back home, discuss Tsujimoto with Marwan, or walk through Kitsilano with Mary at the edge of dark. The PhD is all of it! Thank you all for every part of it.