### 2021-09-25 revision/ack.tex
about to track changes.

%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Acknowledgments}

Attending university long enough to earn a PhD is an unbelievable privilege, and I only had access to it thanks to an incredible amount of guidance and support.

Most of all, to mom Calisa Pierce, dad Jim Pierce, my sisters Kelsey and Kim, and my grandmother Wilma Avis, it's impossible to explain how thankful I am for your guidance. Thanks also to my great aunt Alice who graciously paid for my undergraduate education.

Thanks to my early teachers who put in the extra care to keep me on track, Jonathan Ramey, Linda Ball, Damon Spurlock, Linda Mendez, Linda Afzalirad, Phyllis Adkins, Brandon Willard, Marty Ojeda, and Red Hamlin among many others.

Thanks also to my professors at Southern WV Community College who fostered my way of thinking and provided many of the tools I used for the research in this thesis.
At SWVCC, I'd like to thank especially Mindy Saunders, Tex Wood, Anne Klein, Charles Wood, Don Saunders, and George Trimble. SWVCC was a personal transformation and a great privilege.

At WVU, I'd like to thank Alan Bristow, Tudor Stanescu, and Leo Golubovic in particular. Alan and Tudor, thank you for providing my first opportunities in research. Leo, you are the key person in my development as a researcher.

To my extended cohort at WVU physics, Payne, Megat, Collins, Scott, Dustin, Rice, Wolfe, Samet, Craig, April, Evan, Larry, Robert, Stephen, Rick, Matt, Caleb, John, and all of you others, thanks for the collaboration and the hilarious times. May the lounge live on forever!

To Marwan Hassan, some time ago now I wandered into your office saying I liked rivers and wanted to describe sediment transport with statistical physics. I realize now when I said ``statistical physics" your eyes lit up with visions of Einstein, Nakagawa, Yano, and so many other names I now know well. I was lucky to use the right trigger words.
Marwan, thanks for all of the guidance, for lending me many dozens of books and papers, for the freedom to explore ideas, and most of all for showing me that it's ok (and implicitly expected) to be human in science.

My friends in geography have also been great. Thanks to the whole extended research family: Shawn, Matteo, Conor, Nisreen, Yinlue, Alex, Kyle, Leo, Dave, Tobias, Katie, Maria, Elli, Jiamei, Xingyu, Niannian, and Emma, plus the outsider Eaton-ites, Will, Dave, Rose, and Anya. In particular thanks to Shawn for all of the shared speculations.

Thank you Rui Ferreira and Brett Eaton!
Rui, you welcomed me to Portugal, offered me guidance in the laboratory, and educated me on turbulence and hydraulics. Thanks especially for the thoughtful comments on my thesis.
Brett, you were the first person to welcome me to UBC, and I'll always appreciate your effort to teach me how to write, interpret flume experiments, and understand simplified modeling in light of real world complexity.

To David Furbish, before I started my PhD when I was still a physics kid exclusively, I recall attending your colloquium at UBC with Marwan's invitation. At the end I asked a vague question about whether Langevin equations might be useful for describing sediment transport, and the answer was ``absolutely yes". Now I have a thesis on it. Thanks for the confirmation! And to the Furbish people -- Rachel, Sarah W, Sarah Z, Nakul, Tyler, Erika, and Shawn, thanks sincerely for letting me board the spaceship.

Among UBC Geographers I'd like to thank Nina Hewitt in particular. It was always a pleasure to work for you on your courses, which are so uniquely organized and student-focused. UBC Geography is a much greater department with you in it.

Finally, Mary! My PhD has not been great in terms of work-life balance. Thanks for putting up with my late bedtimes, early wake-ups, rambling science jargon, and skipped social events. It's done!

Completing a PhD looks like a personal achievement, but it's really not.
It's an expression of 30 years of influence and guidance from other people at every stage, my parents, my sisters, my grandmother, and a huge network of teachers and mentors. Thank you all!