### 2021-09-25 revision/app_ch2.tex

%%!TEX root = diss.tex

\chapter{Calculations involved in the dynamical bedload flux model}
\label{sec:appendixAfluc}
\section{Derivation of the master equation for particle position}
\label{sec:appAmaster}
Here I derive the master equation \ref{eq:flippymaster} for the probability distribution of particle position from the Langevin equation \ref{eq:flippylangevin}. This is closely based off on the approach of \citet{Balakrishnan1993}.

The master equation satisfies $P(x,t) = \bra\bra \delta(x-x(t))\ket_\xi \ket_\eta$, where the averages are over all realizations of the two independent noises. It is most convenient to take these averages after taking Fourier transforms and manipulating time derivatives of the resulting equations.

Integrating the Langevin equation \ref{eq:flippylangevin}, substituting the solution in the above definition, and taking the Fourier transform in space gives
\be \tilde{P}(g,t) = \Big\bra  \Big\bra \exp \Big[- i g \int_0^t du [V+\sqrt{2D}\xi(u)]\eta(u) \Big]\Big\ket_\eta \Big\ket_\xi\ee
This can be rearranged as
\be \tilde{P}(g,t) = \Big\bra \exp\Big[i g V \int_0^t du \eta(u)\Big] \Big\bra \exp\Big[ i g \sqrt{2D}\int_0^t du \xi(u) \eta(u)]\Big]\Big\ket_\xi \Big\ket_\eta .\ee

Using the classic identity for an average over white noise of an exponential \citep{Balakrishnan1993,VanKampen2007} gives, after recognizing that the dichotomous noise satisfies $\eta^2 = \eta$ since $\eta$ is either $0$ or $1$,
\be \tilde{P}(g,t) = \Big\bra \exp\Big[(igV-g^2D) \int_0^t du \eta(u) \Big]\Big\ket_\eta. \label{eq:logi}\ee

Now I will take time derivatives in order to apply the so-called Furutsu-Novikov formula to evaluate the ensemble average over dichotomous noise \citep{Shapiro1978}. Applied to the dichotomous noise in Fig. \ref{fig:lislefig}, the Furutsu-Novikov formula is, for an arbitrary functional $F$ of the noise,
\be \pt \bra \eta(t)F[\eta(t)] \ket = \bra \eta \pt F \ket + k \big[\bra \eta \ket \bra F \ket -\bra \eta F \ket \big]  \label{eq:furuti}.\ee

Making the short-hands $G = igV-g^2D$ and $F=\exp G\int_0^t \eta(u)du$ and taking a time derivative of Eq. \ref{eq:logi} gives
\be G^{-1}\pt \tilde{P}(g,t) = \bra \eta F\ket. \label{eq:furut}\ee
Taking a second time derivative opens up the possibility of using the Furutsu-Novikov formula \ref{eq:furuti}: 
\be G^{-1}\pt \tilde{P}(g,t) = \pt \bra \eta F\ket = G\bra \eta F \ket + k\big[ \bra \eta \ket \tilde{P}(g,t) - G^{-1} \pt \tilde{P}(g,t) \big]. \ee
Applying Eq. \ref{eq:furut} finally gives
\be \pt^2 \tilde{P}(g,t)  = (igV-g^2D-k)\pt  \tilde{P} + k_E (igV-g^2D) \tilde{P},\ee
and inverse Fourier transforming provides the master equation \ref{eq:flippymaster}.

\section{Solution for the position probability distribution}
\label{sec:fluccymastersol}
The position probability distribution can be obtained from Eq. \ref{eq:flippymaster} for the initial conditions $P(x,0) = \delta(x)$ and $ \pt P(x,0) = \frac{k_E}{k}\big[D\delta''(x)-V \delta'(x) \big]$ using Fourier transforms in space and Laplace transforms in time. Taking these transforms provides
\be \bar{\tilde{P}}(g,s) = \frac{s + k +\varphi  Dg^2 - i g V \varphi}{s(s+k)+(Dg^2-igV)(s+k_E)},  \label{eq:prev}\ee
where $\varphi = k_D/k$ is the probability a particle starts at rest.

The numerator terms encode the initial conditions. The denominator terms are where the real structure of the solution is contained.
The numerator terms involving $g$ can be expressed as first and second derivatives with respect to $x$. The problem, then is to calculate the inverse Fourier integral of the denominator.
The inverse Fourier transform can be calculated by expanding the denominator of Eq. \ref{eq:prev} in partial fractions and applying the contour integral \citep[e.g.][]{Arfken1985}
\be \int_{-\infty}^\infty \frac{1}{2\pi i} \frac{e^{-i g x}}{g - i c} dg = \sigma(c)\theta(-x\sigma(c))\exp\big(-|c x|\big). \label{eq:contour}\ee
In this equation $\sigma(c) = \text{sgn}(c)$ is a shorthand for the sign function. The above integral can be obtained with the residue theorem by considering different cases for the signs of $x$ and $c$ and applying square contours around the upper or lower-half regions of the complex plane.

Rearranging the governing equation gives
\be \tilde{\hat{P}}(g,s) = \frac{\varphi D g^2 - ig V\varphi  + s + k}{D(s+k_E)}\frac{1}{g^2-i\frac{V}{D}g + \frac{s(s+k)}{D(s+k_E)}}.\ee
The roots of the denominator are at
\be g_\pm = \frac{iV}{2D}\Big[ 1  \pm R \Big],\ee
where
\be R = \sqrt{1 + \frac{4D}{V^2}\frac{s(s+k)}{s+k_E}}.\ee
The partial fractions expansion is therefore
\be \frac{1}{g^2-i\frac{V}{D}g + \frac{s(s+k)}{D(s+k_E)}} = \frac{D}{i V R}\Big[ \frac{1}{g-g_+} - \frac{1}{g-g_-}\Big],\ee
giving 
\begin{multline} \tilde{P}(x,s) = \frac{-\varphi D\px^2 + V\varphi\px + s + k}{VR(s+k_E)} \\ \times \Bigg[ \int \frac{dg}{2\pi i}\frac{e^{-igx}}{g-g_+} - \int \frac{dg}{2\pi i}\frac{e^{-igx}}{g-g_-}\Bigg],\end{multline}
or eventually
\begin{multline}
 \tilde{P}(x,s) = \frac{-\varphi D\px^2 + V\varphi \px + s + k}{VR(s+k_E)} \\ \times \Bigg[ \theta(x) \exp\Big(\frac{Vx}{2D}(1-R)\Big) + \theta(-x)\exp\Big( \frac{Vx}{2D}(1+R)\Big) \Bigg].\end{multline}
This simplifies slightly to
\be \tilde{P}(x,s) = \frac{-\varphi D\px^2 + V\varphi \px + s + k}{VR(s+k_E)}\exp\Big[\frac{Vx}{2D} - \frac{V|x|}{2D}R \Big] \label{eq:laplaceapp}.
\ee

Taking the inverse Laplace transform symbolically provides the desired probability density function
\begin{multline} P(x,t) = \mathcal{L}^{-1}\Bigg\{ \frac{1}{V} \big[-\varphi D\px^2 + V\varphi \px + k + s \big] \\ \times \exp\Big(\frac{Vx}{2D}\Big) \frac{\exp\Big(-\frac{V|x|R}{2D} \Big)}{(s+k_E)R} \Bigg\}(t). \end{multline}
Using the shift property of Laplace transforms \citep[e.g.][]{Arfken1985}, this becomes
\begin{multline} P(x,t) = \frac{1}{V}e^{-k_E t} \mathcal{L}^{-1}\Bigg\{ \big[-\varphi D\px^2 + V \varphi \px + k_D + s\big] \\ \times \exp\Big(\frac{Vx}{2D}\Big) \frac{\exp\Big(\frac{-V|x|R_\ast}{2D}\Big)}{s R_\star} \Bigg\},\end{multline}
where 
\be R_\ast = \sqrt{1 + \frac{4D(k_D-k_E)}{V^2}+ \frac{4D}{V^2}\Big(s-\frac{k_Ek_D}{s}\Big)}. \ee
Using the property that $\mathcal{L}^{-1}\big\{s \tilde{f}\big\} = (\delta(t) + \pt)f(t)$ \citep{Arfken1985} gives
\begin{multline}
 P(x,t) = \frac{1}{V} e^{-k_E t} \big[ -\varphi D\px^2 + V\varphi \px + k_D + \delta(t) + \pt \big] \\ \times \exp\Big[\frac{Vx}{2D}\Big] \mathcal{L}^{-1}\Bigg\{ \frac{-\exp\Big[\frac{-V|x|R_\ast}{2D}\Big]}{sR_\ast} \Bigg\}(t).
\end{multline}
Finally, using the identity
\be \mathcal{L}^{-1} \Big\{ \frac{1}{s} \tilde{g}(s-a/s)\Big\} = \int_0^t \mathcal{I}_0\Big(2\sqrt{au(t-u)}\Big) g(u) du
\ee
from \citet{Bateman1953}, pg. 133 (which can be derived by either u-substitution or series expansion in powers of $a$) provides
\begin{multline}  P(x,t) = \frac{1}{V} e^{-k_E t}  \big[ -\varphi D\px^2 + V\varphi \px + k_D + \delta(t) + \pt \big] \exp\Big[\frac{Vx}{2D}\Big]\\ \times \int_0^t du \mathcal{I}_0\Big( 2 \sqrt{k_Ek_D u(t-u)}\Big) \\ \times \mathcal{L}^{-1}\Bigg\{ \frac{\exp\Big[ \frac{-V|x|}{2D}\sqrt{a+b s}\Big]}{\sqrt{a + b s}}\Bigg\}(u) , \label{eq:magic}\end{multline}
where $a = 1 + 4D(k_E-k_E)/(V^2)$ and $b = 4D/V^2$ are short-hands.

Using the trick of introducing a parameter and integrating over it to remove the square root from the denominator (an integral analogue of "Feynman's trick"), the remaining transform can be evaluated from standard tables \citep{Prudnikov1992a}, eventually giving (after integrating over and setting to $1$ the fake parameter),
\begin{multline} P(x,t) = \big[-\varphi D\px^2 + V\varphi \px + k + \delta(t)+ \pt \big]\int_0^t \mathcal{I}_0\Big( 2 \sqrt{k_Ek_D u(t-u)}\Big) e^{-k_E(t-u)} \\ \times \sqrt{ \frac{1}{4\pi D u}} \exp\Big[-k_D u - \frac{(x-Vu)^2}{4Du}\Big] du, \end{multline}
as reported in Eq. \ref{eq:monster}.

\section{Calculation of the scale-dependent rate function $\Lambda(T)$}
\label{sec:fluxconstant}
The central object required to calculate the sediment flux probability distribution in Eq. \ref{eq:flippyflux} is 
\be \Lambda(T) = \rho \int_0^\infty dx_i \int_0^\infty dx P(x+x_i,T).\ee
This represents the rate at which particles start at $x_i$, anywhere to the left of $x=0$, and manage to cross $x=0$ by time $T$.

Again it is easiest to conduct the necessary calculus after Laplace transforming. This gives 
\be \tilde{\Lambda}(s) = \rho \int_0^\infty dx_i \int_0^\infty dx \tilde{P}(x+x_i,s).\ee
The Laplace transform of the rate function therefore follows after integrating Eq. \ref{eq:laplaceapp} twice.

Noting that $x+x_i$ is always positive,
The first integration gives
\begin{multline} \tilde{\Lambda}(s) = \rho \int_0^\infty dx_i \exp\Big[\frac{V(1-R)x_i}{2D}\Big]\\ 
	\times\Big(\varphi \frac{1-R}{2R(s+k_E)} - \frac{\varphi}{R(s+k_E)} - \frac{2D(s+k)}{V^2R(1-R)(s+k_E)}\Big), \end{multline}
and the second integration provides
\begin{multline}  \tilde{\Lambda}(s) = -\frac{\rho \varphi D}{V R(s+k_E)}  + \frac{2\rho D\varphi}{VR(1-R)(s+k_E)} \\ + \frac{4\rho D^2(s+k)}{V^3R(1-R)^2(s+k_E)}. 
	\label{eq:laplacefluxrate}\end{multline}

Taking the inverse transform, converting the $s$ factor in the numerator of the last term to $\pt + \delta(t)$, and using the shift property gives \citep[e.g.][]{Arfken1985}
\be \Lambda(t) = \rho \mathcal{L}^{-1}\Bigg\{ - \frac{\varphi D}{VR_\star s} + \frac{2D\varphi}{VR_\ast(1-R_\star)s} + \frac{4D^2\big( \cev{\pt} + k\big)}{V^3 R_\star(1-R_\star)^2s}\Bigg\}.\ee
Here the notation $\cev{\pt}$ means the derivative acts from the left on all terms multiplying it (as in $f(t)\cev{\pt} g(t) = \pt [f(t)g(t)]$).

After some work, using Eq. \ref{eq:magic} and tabulated Laplace transform pairs \citep[e.g.][]{Arfken1985,Prudnikov1992a} to perform the inverse Laplace transforms, the net result for the rate constant is
\begin{multline} 
	\Lambda(t) = \rho \int_0^t \mathcal{I}_0\Big(2\sqrt{k_Ek_Du(t-u)}\Big)e^{-k_E(t-u)-k_D u} \\
	\times \Bigg[\sqrt{\frac{D}{\pi u}}\Big([\cev{\partial_t} + k]u-\frac{k_D}{2 k}\Big)e^{-V^2 u/4D}  \\ + \frac{V}{2}\Big([\cev{\partial_t} + k]u -\frac{k_D}{k}\Big) \erfc\Bigg(-\sqrt{\frac{V^2 u}{4D}}\Bigg)\Bigg] du,
\end{multline}
as reported in Eq. \ref{eq:fluxy}.

\label{sec:fluxlimits}
Eq. \ref{eq:fluxy} is difficult to work with. To explore the behavior at extreme values of the observation time $T$, one can apply Tauberian theorems \citep{Weiss1994} to invert the Laplace-transformed rate function of Eq. \ref{eq:laplacefluxrate} at the opposite extreme of $s$.

For example, at short times, expanding Eq. \ref{eq:laplacefluxrate} as $s\rightarrow \infty$ gives
\be \tilde{\Lambda}(s) = \frac{\rho k_E V}{2 k s^2} + \frac{\rho k_E}{k}\sqrt{\frac{D}{4 s^3}} \ee
which inverts to
\be \Lambda(t) \sim \frac{\rho k_E V T}{2 k} + \frac{\rho k_E}{k}\sqrt{\frac{D T}{\pi}},\ee
giving the small $T$ behavior.
This has two scaling limits within it. Provided that $T \ll 4D/V^2/\pi < 2 D/V^2$, the scaling goes as $\Lambda(T) \sim T^{-1/2}$. But if $T\gg 4D/V^2/\pi$, it goes as $\Lambda(t) \sim T$.
For large times, taking $s\rightarrow 0 $ gives
\be \tilde{\Lambda}(s) = \frac{\rho k_E V }{ks^2}, \ee
and this inverts to $\Lambda(T) = \rho k_E V T/k$ as the long time solution, providing a mean flux equivalent to the Einstein model (Sec. \ref{sec:lisle}).
These limits are summarized in Eq. \ref{eq:fluccolimitcases}.