### 2021-09-25 revision/ch4downDiff.tex

%%!TEX root = diss.tex

\chapter{Burial-induced three-range diffusion in sediment transport}
\label{ch:downDiff}

Many environmental problems including channel morphology \citep{Hassan2017}, contaminant transport \citep{Macklin2006}, and aquatic habitat restoration \citep{Gaeuman2017} rely on our ability to predict the diffusion characteristics of coarse sediment tracers in river channels.
Diffusion is quantified by the time dependence of the positional variance $\sigma_x^2$ of a group of tracers.
With the scaling $\sigma_x^2 \propto t$, the diffusion is said to be normal, since this is found in the classic problems \citep{Einstein1905}.
However, with the scaling $\sigma_x^2 \propto t^\gamma$ with $\gamma \neq 1$, the diffusion is said to be anomalous \citep{Sokolov2012}, with $\gamma>1$ defining super-diffusion and $\gamma<1$ defining sub-diffusion \citep{Metzler2000}.
\citet{Einstein1937} developed one of the earliest models of bedload diffusion to describe a series of flume experiments \citep{Ettema2004}.
Interpreting individual bedload trajectories as a sequence of random steps and rests, Einstein originally concluded that a group of bedload tracers undergoes normal diffusion.

More recently, Nikora et al. realized coarse sediment tracers can show either normal or anomalous diffusion depending on the length of time they have been tracked \citep{Nikora2001a,Nikora2002}.
From numerical simulations and experimental data, Nikora et al. discerned ``at least three'' scaling ranges $\sigma_x^2 \propto t^\gamma$ as the observation time increased.
They associated the first range with ``local'' timescales less than the interval between subsequent collisions of moving grains with the bed, the second with ``intermediate'' timescales less than the interval between successive resting periods of grains, and the third with ``global'' timescales composed of many intermediate timescales.
Nikora et al. proposed super-diffusion in the local range, anomalous or normal diffusion in the intermediate range, and sub-diffusion in the global range.
They attributed these ranges to ``differences in the physical processes which govern the local, intermediate, and global trajectories'' of grains \citep{Nikora2001a}, and they called for a physically based model to explain the diffusion characteristics \citep{Nikora2002}.

Experiments support the Nikora et al. conclusion of multiple scaling ranges \citep{Martin2012,Fathel2016}, but they do not provide consensus on the expected number of ranges or their scaling properties.
This lack of consensus probably stems from resolution issues.
For example, experiments have tracked only moving grains, resolving the local range \citep{Furbish2012,Furbish2017,Fathel2016}; grains resting on the bed surface between movements, resolving the intermediate range \citep{Einstein1937,Yano1969,Nakagawa1976}; grains either moving or resting on the bed surface, likely resolving local and intermediate ranges \citep{Martin2012}; or grains resting on the surface after floods, likely resolving the global range \citep{Phillips2013,Bradley2017}. 
At long timescales, a significant fraction of tracers become buried under the bed surface \citep{Hassan1991,Hassan2013,Ferguson2002a,Haschenburger2013,Papangelakis2016}, meaning burial dominates long term diffusion characteristics \citep{Bradley2017,Martin2014,Voepel2013}, possibly at global or even longer ``geomorphic'' timescales \citep{Hassan2017} than Nikora et al. originally considered.
As a result, three diffusion ranges can be identified by patching together multiple data-sets \citep{Zhang2012,Nikora2002}, but they are not resolved by any one data-set.

Newtonian bedload trajectory models also show multiple diffusion ranges, although they also do not provide consensus on the expected number of ranges or their scaling properties. 
The majority of these models predict two ranges of diffusion (local and intermediate) without predicting a global range.
Among these, \citet{Nikora2001a} used synthetic turbulence \citep{Kraichnan1970} with a discrete element method for the granular phase \citep{Cundall1979}; \citet{Bialik2012} used synthetic turbulence with a random collision model \citep{Sekine1992}; and \citet{Fan2016} used a Langevin equation with probabilistic rests.
To my knowledge, only \citet{Bialik2015} have claimed to capture all three ranges from a Newtonian approach.
They incorporated a second resting mechanism into their earlier model \citep{Bialik2012}, implicitly suggesting that three diffusion ranges could result from two distinct timescales of sediment rest.
However, Newtonian approaches have not evaluated the effect of sediment burial on tracer diffusion, probably due to the long simulation timescales required.

Random walk bedload diffusion models constructed in the spirit of \citet{Einstein1937} provide an alternative to the Newtonian approach and can include a second timescale of rest by incorporating sediment burial.
Einstein originally modeled bedload trajectories as instantaneous steps interrupted by durations of rest lying on statistical distributions \citep{Hassan1991}, but this generates only one range of normal diffusion \citep{Einstein1937,Hubbell1964,Nakagawa1976}.
Recently, researchers have generalized Einstein's model in a few different ways to describe multiple diffusion ranges.
\citet{Lisle1998} and \citet{Lajeunesse2017} promoted Einstein's instantaneous steps to motion intervals with random durations and a constant velocity, providing two diffusion ranges -- local and intermediate.
\citet{Wu2019} retained Einstein's instantaneous steps but included the possibility that grains can become permanently buried as they rest on the bed, also providing two diffusion ranges -- intermediate and global. 
These earlier works suggest the minimal required components to model three bedload diffusion ranges: (1) exchange between motion and rest intervals and (2) the sediment burial process.

In this study, I incorporate these two components into Einstein's original approach to describe three diffusion ranges with a physically based model, as called for by \citet{Nikora2002}.
Einstein was a giant in river geophysics and fostered an entire paradigm of research leveraging and generalizing his stochastic methods \citep{Hubbell1964, Yano1969, Yang1971, Gordon1972, Nakagawa1976,Paintal1971}.
Einstein's model can be viewed as a pioneering application of the continuous time random walk (CTRW) developed by \citet{Montroll1965} in condensed matter physics to describe the diffusion of charge carriers in solids.
To incorporate motion intervals and sediment burial, I utilize the multi-state CTRW formalism developed by \citet{Weiss1976, Weiss1994} that extends the CTRW of \citet{Montroll1965}.
Below, I develop and solve the model in Sec. \ref{sec:model}. Then, I discuss the predictions of this model, present its implications for local, intermediate, and global ranges of bedload diffusion, and suggest next steps for bedload diffusion research in Secs. \ref{sec:discussion} and \ref{sec:conclusion}.

\section{Bedload trajectories as a multi-state random walk}
\label{sec:model}
\subsection{Assumptions of the burial model}
\label{sec:assumptions}
Particle trajectories are formulated as a three-state random walk where the states are motion, surface rest, and burial. These states are labeled as $i=2$ (motion), $i=1$ (rest), and $i=0$ (burial).
The target is the probability distribution $p(x,t)$ to find a grain at position $x$ and time $t$ if it is known to have started with the initial distribution $p(x,0)=\delta(x)$.
Times spent moving or resting on the surface are characterized by exponential distributions $\psi_M(t)=k_2e^{-k_2 t}$ and $\psi_1(t) = k_1e^{-k_1t}$, since numerous experiments show thin-tailed distributions for these quantities \citep{Fathel2015,Roseberry2012,Einstein1937,Ancey2006,Martin2012}. The end results will not be contingent on the specific distributions chosen, since all thin-tailed distributions provide similar diffusion characteristics in random walks \citep{Weiss1994,Weeks1998}.
Grains in motion are considered to have a characteristic velocity $v$ \citep{Lisle1998,Lajeunesse2017}, and burial is modeled as long lasting enough to be effectively permanent \citep{Wu2019}, with grains resting on the surface having a probability per unit time $\kappa$ to become buried.
This means $\Phi(t) = e^{-\kappa t}$ represents the probability that a grain is not buried after resting for a time $t$, while $1-\Phi(t)$ represents the probability that it is buried.
The initial conditions are specified with probabilities $\theta_1$ and $\theta_2$ to be in rest or motion at $t=0$. Normalization requires $\theta_1+\theta_2=1$.

\subsection{Governing equations}
With these assumptions, the governing equations for the set of probabilities $\om_{ij}(x,t)$ that a transition occurs from state $i$ to state $j$ at position $x$ and time $t$ can be derived using the statistical physics approach to multi-state random walks \citep{Weiss1994,Schmidt2007,Weeks1998}.
Denoting by $g_{ij}(x,t)$ the probability for a particle to move a distance $x$ in a time $t$ within the state $i$ before it transitions to the state $j$, the transition probabilities $\om_{ij}(x,t)$ sum over all possible paths to the state $i$ from previous locations and times:
\be \om_{ij}(x,t) = \theta_i g_{ij}(x,t) + \sum_{k=0}^2 \int_0^x dx' \int_0^t dt' \om_{ki}(x',t')g_{ij}(x-x',t-t').\label{eq:g1}\ee
Defining another set of probabilities $G_i(x,t)$ that a particle moves by a distance $x$ in a time $t$ within the state $i$ and possibly remains within the state, a similar sums over paths for the probabilities to be in the state $i$ at $x$, $t$ produces: 
\be p_i(x,t) = \theta_i G_i(x,t) + \sum_{k=0}^2 \int_0^x dx' \int_0^t dt' \om_{ki}(x',t')G_i(x-x',t-t').\label{eq:g2}\ee
Finally, the overall probability to be at position $x$ at time $t$ is
\be p(x,t) = \sum_{k=0}^2 p_k(x,t). \ee
This joint density is completely determined from the solutions of Eqs. \ref{eq:g1} and \ref{eq:g2} given specification of the distributions $g_{ij}$ and $G_i$.

\subsection{Joint probability distribution of particle position with burial}
\label{sec:solution}

These distributions can be constructed from the assumptions described in Sec. \ref{sec:assumptions}.
Since particles resting on the bed surface become buried in a time $t$ with probability $\Phi(t)$, and resting times are distributed as $\psi_1(t)$, these assumptions obtain $g_{12}(x,t) = \delta(x)k_1e^{-k_1t}e^{-\kappa t}$ and $g_{10}(x,t) = \delta(x) k_1 e^{-k_1 t}(1-e^{-\kappa t})$. Since motions have velocity $v$ for times distributed as $\psi_2(t)$, the assumptions provide $g_{21}(x,t) = \delta(x-vt)k_2e^{-k_2 t}$.
Since burial is quasi-permanent, all other $g_{ij} = 0$.
The $G_i$ are constructed in the same way except using the cumulative probabilities $\int_t^\infty dt'\psi_i(t) = e^{-k_i t}$, since these characterize motions and rests that are ongoing \citep{Weiss1994}.
Using the cumulative probabilities provides $G_1(x,t) = \delta(x)e^{-k_1t}$ and $G_2(x,t) = \delta(x-vt)e^{-k_2 t}$.

Eqs. \ref{eq:g1} and \ref{eq:g2} with these $g_{ij}$ and $G_i$ are solved using Laplace transforms in space and time $(x,t \rightarrow 
\eta,s).$ This method, similar to \citet{Weeks1998}, unravels the convolution structure of these equations, eventually producing
\be \tilde{p}(\eta,s) = \frac{1}{s}\frac{(s+\kappa + k')s  + \theta_1(s+\kappa )\eta v+ \kappa k_2}{(s+\kappa+k_1)\eta v+(s+\kappa+k')s + \kappa k_2}, \label{eq:diffnicedist}\ee
where $k' = k_1 + k_2$. As explained in appendix Sec. \ref{sec:appendixA}, inverting this result using known Laplace transforms \citep{Prudnikov1992a,Arfken1985} obtains
\begin{align}
	\begin{split}
		p(x,t) = \theta_1&\Big[1-\frac{k_1}{\kappa+k_1}\Big(1-e^{-(\kappa+k_1)t}\Big)\Big]\delta(x) \\ &+ \frac{1}{v}e^{-\Omega \tau - \xi}\Big(\theta_1\Big[k_1\mathcal{I}_0\big(2\sqrt{\xi\tau}\big) + k_2\sqrt{\frac{\tau}{\xi}}\mathcal{I}_1\big(2\sqrt{\xi\tau}\big)\Big] \\ 
		&+ \theta_2\Big[k_1\delta(\tau) + k_2 \mathcal{I}_0\big(2\sqrt{\xi\tau}\big)+k_1 \sqrt{\frac{\xi}{\tau}}\mathcal{I}_1\big(2\sqrt{\xi\tau}\big)\Big]\Big) \\
		&+ \frac{1}{v}\frac{\kappa k_2}{\kappa + k_1}e^{-\kappa \xi/(\kappa + k_1)}\Big[(\theta_1/\Omega)\mathcal{Q}_2(\xi/\Omega,\Omega\tau) + \theta_2 \mathcal{Q}_1(\xi/\Omega,\Omega\tau)\Big]
		\label{eq:pdf}
	\end{split}
\end{align}
for the joint distribution that a tracer is found at position $x$ at time $t$.
This result generalizes the earlier results of \citet{Lisle1998} and \citet{Einstein1937} to include sediment burial.
This equation uses the shorthand notations $\xi = k_2 x/v$, $\tau = k_1(t-x/v)$, and $\Omega = (\kappa+k_1)/k_1$ \citep[cf.][]{Lisle1998}. The $\mathcal{I}_\nu$ are modified Bessel functions of the first kind and the $\mathcal{Q}_\mu$ are generalized Marcum Q-functions defined by $\mathcal{Q}_\mu(x,y) = \int_0^y e^{-z-x}(z/x)^{(\mu-1)/2}\mathcal{I}_{\mu-1}(2\sqrt{xz})dz $ and originally devised for radar detection theory \citep{Marcum1960,Temme1996}. 
The Marcum Q-functions derive from the burial process.
Because resting grains can become buried with an exponential probability, while the probability that particles rest follows a modified Bessel distribution \citep{Einstein1937,Lisle1998}, evaluating the probability that particles rest and become buried generates the Q-function convolution structure.

\begin{figure}[!htbp]
	\includegraphics[width=\linewidth,keepaspectratio]{./figures/ch4/pdf-plot.png}
	\caption{Joint distributions for a grain to be at position $x$ at time $t$ are displayed for the choice $k_1=0.1$, $k_2=1.0$, $v=2.0$. Grains are considered initially at rest ($\theta_1=1$, $\theta_2=0$). The solid lines are the analytical distribution in Eq. \ref{eq:pdf}), while the points are numerically simulated, showing the correctness of the derivations. Colors pertain to different times. Units are unspecified, since the aim is to demonstrate the general characteristics of $p(x,t)$. Panel (a) shows the case $\kappa=0$ -- no burial. In this case, the joint distribution tends toward Gaussian at large times \citep{Einstein1937,Lisle1998}. Panel (b) shows the case when grains have rate $\kappa = 0.01$ to become buried while resting. Because of burial, the joint distribution tends toward a more uniform distribution than Gaussian.
		\label{fig:diffpdfs}}
\end{figure}

Fig. \ref{fig:diffpdfs} depicts the distribution Eq. \ref{eq:pdf} alongside simulations generated by a direct method based on evaluating the cumulative transition probabilities between states on a small time-step \citep{Barik2006}.
When grains do not become buried, as in panel (a) of Fig. \ref{fig:diffpdfs}, the distribution becomes Gaussian-like at relatively large observation times, exemplifying normal diffusion and satisfying the central limit theorem.
When grains become buried, as in panel (b) of Fig. \ref{fig:diffpdfs},  the Q-function terms prevent the distribution from approaching a Gaussian at large timescales, exemplifying anomalous diffusion \citep{Weeks1998} and violating the central limit theorem \citep{Metzler2000,Schumer2009}.

\subsection{Downstream diffusion}

To obtain an analytical formula for tracers diffusing downstream while they gradually become buried, the first two moments of position are derived by taking derivatives with respect to $\eta$ of the Laplace space distribution (Eq. \ref{eq:diffnicedist}) using an approach similar to \citet{Shlesinger1974} and \citet{Weeks1998}. These moments produce the positional variance $\sigma_x^2 = \bra x^2\ket - \bra x \ket^2$. 
The first two moments are
\begin{align}
	\bra x(t) \ket &= A_1 e^{(b-a)t}+B_1e^{-(a+b)t}+C_1, \label{eq:mean}\\
	\bra x^2(t) \ket &= A_2(t)e^{(b-a)t}+B_2(t)e^{-(a+b)t}+C_2, \label{eq:second}
\end{align}
so the variance is 
\be \sigma_x^2(t) = A(t)e^{(b-a)t} + B(t)e^{-(a+b)t} + C(t). \label{eq:var}\ee
In these equations, $a = (\kappa + k_1+k_2)/2$ and $b = \sqrt{a^2-\kappa k_2}$ are effective rates having dimensions of inverse time, while the $A$, $B$, and $C$ factors are provided in Tab. \ref{table:params}.

%\\
%		&\hspace{1cm}
\begin{table}[!h]
	\centering
	%\begin{wraptable}{r}{width=5.5cm}
	\caption{Abbreviations used in the expressions of the mean (Eq. \ref{eq:mean}), second moment (Eq. \ref{eq:second}) and variance (Eq. \ref{eq:var}) of bedload tracers.}
	\label{table:params}
	\small
	\begin{tabular}{c}
		\toprule
		$\begin{aligned}[t]
			&A_1 = \frac{v}{2b}\big[\theta_2+\frac{k_1+\theta_2\kappa}{b-a}\big] \\
			&B_1 = -\frac{v}{2b}\big[\theta_2-\frac{k_1+\theta_2 \kappa}{a+b}\big] \\
			&C_1 =  -\frac{v}{2b}\big[\frac{k_1+\theta_2 \kappa}{b-a}+\frac{k_1+\theta_2 \kappa}{a+b}\big]\\
			&A_2(t) = \frac{v^2}{2b^3}\Big[(bt-1)[k_1+\theta_2(2\kappa + k_1 + b-a)]+\theta_2b 
			+ \frac{(\kappa+k_1)(\theta_2\kappa+k_1)}{(b-a)^2}[(bt-1)(b-a)-b]\Big]\\
			&B_2(t) = \frac{v^2}{2b^3}\Big[(bt+1)[k_1 + \theta_2(2\kappa+k_1-a-b)]+\theta_2b
			-\frac{(\kappa+k_1)(\theta_2\kappa+k_1)}{(a+b)^2}[(bt+1)(a+b)+b]\Big]\\
			&C_2 = \frac{v^2}{2b^3}(\kappa+k_1)(\theta_2 \kappa + k_1)\Big[\frac{2b-a}{(b-a)^2}+\frac{a+2b}{(a+b)^2}\Big]\\
			&A(t) = A_2(t)-2A_1C_1 - A_1^2\exp[(b-a)t]\\
			&B(t) = B_2(t)-2B_1C_1 - B_1^2\exp[-(a+b)t]\\
			&C(t) = C_2-C_1^2-2A_1B_1\exp[-2at]\\			
		\end{aligned}$\\
		\bottomrule
	\end{tabular}
	%\end{wraptable}
	\vspace{-0.5cm}
\end{table}
The positional variance Eq. \ref{eq:var} is plotted in Fig. \ref{fig:diffvar} for conditions $\theta_1=1$ and $k_2\gg k_1 \gg \kappa$.
The notation ``$\gg$'' is interpreted in this context to mean ``of at least an order of magnitude greater''.
These conditions are most relevant to tracers in gravel-bed rivers, since they represent that grains are initially at rest \citep{Hassan1991,Wu2019}, motions are typically much shorter than rests \citep{Einstein1937,Hubbell1964}, and burial requires a much longer time than typical rests  \citep{Ferguson2002,Hassan1994,Haschenburger2013}.
Figure \ref{fig:diffvar} demonstrates that under these conditions the variance (Eq. \ref{eq:var}) shows three diffusion ranges with approximate power law scaling ($\sigma_x^2 \propto t^\gamma$) that are identified as the local, intermediate, and global ranges proposed by Nikora et al., followed by a fourth range of no diffusion ($\sigma_x^2 = \text{const}$) stemming from the burial of all tracers. 
Following \citet{Hassan2017} I suggest to call the fourth range "geomorphic", since any further transport in this range can occur only if scour re-exposes buried grains to the flow \citep{Nakagawa1980,Voepel2013,Martin2014,Wu2019a}.

\begin{figure}[!htbp]	
	\includegraphics[width=\linewidth,keepaspectratio]{./figures/ch4/diffusion.png}
	\caption{Panel (a) sketches conceptual trajectories of grains, while panel (b) depicts Eq. \ref{eq:var} with mean motion time $1.5$ s, rest time $30.0$ s, and velocity $0.1$ m/s. The burial timescale is $7200$ s (two hours), and grains start from rest ($\theta_1=1$). The solid line is Eq. \ref{eq:var} and points are numerically simulated. Panel (b) demonstrates four distinct scaling ranges of $\sigma_x^2$: local, intermediate, global, and geomorphic. The first three are diffusive. Crossover times $\tau_L$, $\tau_I$, and $\tau_G$ divide the ranges. Slope keys demonstrate the scaling $\sigma_x^2 \propto t^\gamma$ in each range. Panel (a) demonstrates that different mixtures of motion, rest, and burial states generate the ranges. At local timescales, grains usually either rest or move; at intermediate timescales, they transition between rest and motion; at global timescales, they transition between rest, motion, and burial; and at geomorphic timescales, all grains become buried.
	}
	\label{fig:diffvar}
\end{figure}

\subsection{Diffusion exponents and three-range scaling}
\label{sec:amidonenow}

Two limiting cases of Eq. \ref{eq:var} provide the scaling exponents $\gamma$ of the diffusion $\sigma_x^2 \propto t^\gamma$ in each range. Limit (1) represents times so short that a negligible amount of sediment burial has occurred, $t\ll 1/\kappa$, while limit (2) represents times so long motion intervals appear as instantaneous steps of mean length $l=v/k_2$, $1/k_2 \rightarrow 0$ while $v/k_2 = \text{constant}$.
Limit (1) provides local exponent $2 \leq \gamma \leq 3$ depending on the initial conditions $\theta_i$, and intermediate exponent $\gamma=1$.
If grains start in motion or rest exclusively, meaning one $\theta_i = 0$, the local exponent is $\gamma=3$, while if grains start in a mixture of motion and rest states, meaning neither $\theta_i$ is zero, the local exponent is $\gamma=2$.
Limit (2) provides global exponent $1 \leq \gamma \leq 3$ depending on the relative importance of $\kappa$ and $k_1$.
The extreme $k_1/\kappa \ll 1$ produces $\gamma=1$ in the global range, while the opposite extreme $k_1/\kappa \rightarrow \infty$ produces $\gamma=3$.
To summarize, when $k_2\gg k_1 \gg \kappa$ so all three diffusion ranges exist, Eq. \ref{eq:var} implies:
\begin{enumerate}
	\item local range super-diffusion with $2<\gamma<3$ depending on whether grains start from purely motion or rest ($\gamma=3$) or from a mixture of both states ($\gamma=2$),
	\item intermediate range normal diffusion $\gamma=1$ independent of model parameters, and
	\item global range super-diffusion $1<\gamma<3$ depending on whether burial happens relatively slowly ($\gamma \rightarrow 1$) or quickly ($\gamma \rightarrow 3$) compared to surface resting times.
\end{enumerate}
Finally, the burial of all tracers generates a geomorphic range of no diffusion.

\section{Discussion}
\label{sec:discussion}

\subsection{Local and intermediate ranges with comparison to earlier work}

This chapter has extended \citet{Einstein1937} by including motion, rest, and burial processes in a multi-state random walk \citep{Weiss1994,Weeks1998} to demonstrate that a group of bedload tracers moving downstream while gradually becoming buried will generate a super-diffusive local range \citep{Martin2012,Fathel2016,Witz2018}, a normal-diffusive intermediate range \citep{Nakagawa1976,Yano1969}, and a super-diffusive global range \citep{Bradley2017, Bradley2010} before the diffusion eventually terminates in a geomorphic range \citep{Hassan2017}.
\citet{Nikora2002} highlighted the need for such a physical description, although they suggested to use a two-state random walk between motion and rest states with heavy-tailed resting times, and they did not discuss sediment burial.
However, other works have demonstrated that a two-state walk with heavy-tailed rests provides two diffusion ranges -- not three \citep{Weeks1996,Fan2016}, and although heavy-tailed resting times have been documented for surface particles \citep{Liu2019,Fraccarollo2019}, they are more often associated with buried particles \citep{Martin2012,Martin2014,Voepel2013,Olinde2015,Pelosi2016, Pierce2020a}, while surface particles retain light-tailed resting times \citep{Einstein1937,Yano1969,Ancey2006,Nakagawa1976}.
Accordingly, I developed a random walk model of bedload trajectories with light-tailed surface resting times that incorporates sediment burial.

The local and intermediate range diffusion characteristics resulting from this model correspond closely to the original Nikora et al. concepts, while the global range has a different origin than Nikora et al. envisioned.
\citet{Nikora2001a} explained that local diffusion results from the non-fractal (smooth) characteristics of bedload trajectories between subsequent interactions with the bed,  while intermediate diffusion results from the fractal (rough) characteristics of bedload trajectories after many interactions with the bed.
This chapter represents these conclusions: non-fractal (and super-diffusive) bedload trajectories exist on timescales short enough that each grain is either resting or moving, while fractal (and normal-diffusive) bedload trajectories exist on timescales when grains are actively switching between motion and rest states.
I conclude that local and intermediate ranges stem from the interplay between motion and rest timescales, as demonstrated by earlier two-state random walk models \citep{Lisle1998,Lajeunesse2017} and by all Newtonian models that develop sequences of motions and rests \citep{Nikora2001a, Bialik2012}, even those including heavy-tailed rests \citep{Fan2016}.

\subsection{Global and geomorphic ranges with next steps for research}

Nikora et al. explained that divergent resting times generate a sub-diffusive global range.
However, studies have demonstrated that divergent resting times can generate super-diffusion in asymmetric random walks \citep{Weeks1996,Weeks1998}, and both experiments \citep{Bradley2017,Bradley2010} and models \citep{Pelosi2016,Wu2019,Wu2019a} of bedload tracers undergoing burial have demonstrated global range super-diffusion.
While my own results also show global range super-diffusion, they do not necessarily refute the Nikora et al. conclusion of sub-diffusion at long timescales.
I assumed sediment burial was a permanent condition which developed a non-diffusive geomorphic range.
In actuality, burial is a temporary condition, because bed scour can exhume buried sediment back into transport \citep{Wu2019a}, probably after heavy-tailed intervals \citep{Voepel2013,Martin2014}.
A generalization of the model in this chapter to include heavy-tailed timescales between burial and exhumation might develop four ranges of diffusion, where the long-time decay of the exhumation time distribution would dictate the geomorphic range diffusion characteristics as depicted in Fig. \ref{fig:diffvar}.
If cumulative exhumation times decay faster than $T^{-1/2}$, as suggested in Ch. \ref{ch:ch3}, other equilibrium transport models \citep{Voepel2013, Martin2014}, and laboratory experiments \citep{Martin2014,Martin2012}, the geomorphic range is expected to be super-diffusive \citep{Weeks1998}.
However, if they decay slower than $T^{-1/2}$, as implicitly suggested by the data of \citet{Olinde2015}, the geomorphic range is expected to be genuinely sub-diffusive \citep{Weeks1998}, leaving Nikora et al. with the final word on long-time sub-diffusion.

The analytical solution of bedload diffusion in Eq. \ref{eq:var} reduces exactly to the analytical solutions of the \citet{Lisle1998} and \citet{Lajeunesse2017} models in the limit without burial ($\kappa \rightarrow 0$), the \citet{Wu2019} model in the limit of instantaneous steps ($k_2 \rightarrow \infty$ and $l = v/k_2$), and the original \citet{Einstein1937} model in the limit of instantaneous steps without burial.
These reductions demonstrate that the majority of recent bedload diffusion models, whether developed from Exner-type equations \citep{Wu2019,Pelosi2014,Pelosi2016} or advection-diffusion equations \citep{Lisle1998,Lajeunesse2017}, can be viewed equivalently as continuous-time random walks applied to individual bedload trajectories.
Within random walk theory, sophisticated descriptions of transport with variable velocities \citep{Zaburdaev2008,Masoliver1994}, correlated motions \citep{Escaff2018,Vicsek2012a}, and anomalous diffusion \citep{Masoliver2016,Fa2014,Metzler2014} have been developed.
Meanwhile, in bedload transport research, variable velocities \citep{Lajeunesse2010,Furbish2012,Heyman2016}, correlated motions \citep{Heyman2014,Lee2018,Saletti2020}, and anomalous diffusion \citep{Fathel2016,Bradley2017,Schumer2009} constitute open research issues.
I believe further developing the linkage between existing bedload models and random walk concepts could rapidly progress our understanding.

\section{Summary}
\label{sec:conclusion}
In this chapter, I developed a random walk model to describe sediment tracers transporting through a river channel as they gradually become buried, providing a physical description of the local, intermediate, and global diffusion ranges identified by \citet{Nikora2002}.
Pushing their ideas somewhat further, I followed \citet{Hassan2017} to propose a geomorphic range describing particle diffusion characteristics at timescales larger than the global range when burial and exhumation both moderate downstream transport.
At base level, this work demonstrates that (1) durations of sediment motions, (2) durations of sediment rest, and (3) the sediment burial process are sufficient to develop three diffusion ranges that terminate when all tracers become buried.
A next step is to incorporate exhumation to better understand the geomorphic range.
Ultimately, the multi-state random walk formalism used in this chapter has been demonstrated to implicitly underlie most existing bedload diffusion models.
The random walk formalism is an alternative perspective on the Langevin approaches used elsewhere in this thesis, and it provides a powerful tool for researchers targeting landscape-scale understanding from statistical concepts of the underlying grain-scale dynamics.