### 2021-09-25 revision/pref.tex

%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

\chapter{Preface}

This thesis is entirely the original research of Kevin Pierce, including all figures, writing, and calculations.
The supervisory committee and especially the research supervisor Marwan Hassan provided guidance in the research.
Suggestions from Michael Church, Rui Ferreira, Roger Beckie, and Peter Ashmore improved the clarity of the writing.

The thesis includes two published works, Chapters \ref{ch:ch3} and \ref{ch:downDiff}, available in \textit{Journal of Geophysical Research: Earth Surface} and \textit{Geophysical Research Letters} respectively, with Marwan Hassan as coauthor \citep{Pierce2020a,Pierce2020b}. Chapters \ref{ch:flux} and \ref{ch:langevin} will be submitted for publication after acceptance of the thesis.