### 2021-09-28 revision/laySum.tex

%% The following is a directive for TeXShop to indicate the main file
%%!TEX root = diss.tex

%% https://www.grad.ubc.ca/current-students/dissertation-thesis-preparation/preliminary-pages
%% 
%% LAY SUMMARY Effective May 2017, all theses and dissertations must
%% include a lay summary.  The lay or public summary explains the key
%% goals and contributions of the research/scholarly work in terms that
%% can be understood by the general public. It must not exceed 150
%% words in length.

\chapter{Lay Summary}

Predicting the flow rates of gravel through river channels is important to manage environmental issues involving rivers.
Most approaches to predict gravel flow consider gravel as if it were a heavy fluid.
These descriptions oversimplify the problem, since gravel flow is actually composed of individual grains bouncing and rolling downstream, which is hardly like a fluid.

In this thesis, I consider gravel flow as the result of individual grains. Using physics and probability, I calculate how fast individual grains move downstream, then I demonstrate how to use their downstream movement paths to predict overall gravel flow rates.
This research produces a more realistic description of gravel movement in rivers which helps us understand how rivers change through time.